Local Sensitivity Analysis of a Supercritical Extraction Model

### 2025-01-11 Sections/Sensitivity_Pressure.tex
11.01

	{\color{blue} As discussed in Chapter \ref{CH:Governing_equations_chapter}, at the low velocity, the fluid can be treated as incompressible, which results in the instant propagation} of pressure throughout the system, enabling a single pressure value to be considered for the entire system. In response to a pressure change, the energy equation experiences a simultaneous deviation across the entire spatial domain. This pressure change impacts the fluid temperature within the computational domain, while boundary values are constrained by conditions specified at the domain's extremes. Dirichlet boundary conditions impose a fixed temperature value at the inlet, creating a thermal gradient that propagates through the system. Alternatively, the Neumann boundary conditions specify a heat flux at the extremes. The zero Neumann boundary conditions are applied to ensure an uniform response of temperatures at the inlet, outlet, and within the extractor to the pressure change.

	Figure \ref{fig:Sensitivty_P_CS} illustrates the sensitivity of solute concentration in the solid phase to pressure changes. As discussed in Chapter \ref{CH: Continuity}, the velocity of a fluid is inversely proportional to its density, indicating that increased pressure reduces the fluid velocity. This results in an extended residence time, allowing for longer interaction between the solute and solvent.	Initially, the extraction process operates in the kinetic-controlled regime, where the concentration gradient is high, and solute solubility is the limiting factor. As noted by \citet{Sliczniuk2024}, the system is considered far from saturation, which explains the low initial system response. This low initial response was also observed in the sensitivity analysis by \citet{Fiori_2007}. The system's response becomes more pronounced as the concentration gradient decreases and the extraction shifts from the kinetic-controlled to the diffusion-controlled regime. {\color{blue}In Figure \ref{fig:Sensitivty_P_CS}, all $\frac{dc_s}{dP}$ values are negative, which indicates a faster mass transfer} from the solid phase, corresponding to enhanced mass transfer.  Over time, as the amount of solute decreases, it becomes a limiting factor, reducing the effect of pressure changes on the system. Eventually, the sensitivities approach zero asymptotically as the solute is washed out from the bed.

@@

	Figure \ref{fig:Sensitivty_P_CF} illustrates the sensitivity of solute concentration in the fluid phase to pressure changes. Compared to Figure \ref{fig:Sensitivty_P_CS}, the dynamic behaviour of fluid phase sensitivities is more apparent. {\color{blue}Due to advection, the sensitivities related to the fluid phase moves across the system in the direction of flow}. Initially, the system response is low despite the pressure increase improving mass transfer, reflecting the previously discussed idle period. As the process continues, the sensitivities increase, indicating faster mass transfer from the solid phase. The corresponding improvement in the mass transfer from the solid particles to the fluid is represented by the positive values in Figure \ref{fig:Sensitivty_P_CF}. When the solute in the solid phase becomes a limiting factor, the extraction rate slows, and sensitivities eventually approaching zero asymptotically.

>>>>

	As discussed in Chapter \ref{CH:Governing_equations_chapter}, at the low velocity, the fluid can be treated as incompressible, which results in the instant propagation of pressure throughout the system, enabling a single pressure value to be considered for the entire system. In response to a pressure change, the energy equation experiences a simultaneous deviation across the entire spatial domain. This pressure change impacts the fluid temperature within the computational domain, while boundary values are constrained by conditions specified at the domain's extremes. Dirichlet boundary conditions impose a fixed temperature value at the inlet, creating a thermal gradient that propagates through the system. Alternatively, the Neumann boundary conditions specify a heat flux at the extremes. The zero Neumann boundary conditions are applied to ensure an uniform response of temperatures at the inlet, outlet, and within the extractor to the pressure change.

	Figure \ref{fig:Sensitivty_P_CS} illustrates the sensitivity of solute concentration in the solid phase to pressure changes. As discussed in Chapter \ref{CH: Continuity}, the velocity of a fluid is inversely proportional to its density, indicating that increased pressure reduces the fluid velocity. This results in an extended residence time, allowing for longer interaction between the solute and solvent.	Initially, the extraction process operates in the kinetic-controlled regime, where the concentration gradient is high, and solute solubility is the limiting factor. As noted by \citet{Sliczniuk2024}, the system is considered far from saturation, which explains the low initial system response. This low initial response was also observed in the sensitivity analysis by \citet{Fiori_2007}. The system's response becomes more pronounced as the concentration gradient decreases and the extraction shifts from the kinetic-controlled to the diffusion-controlled regime. In Figure \ref{fig:Sensitivty_P_CS}, all $\frac{dc_s}{dP}$ values are negative, which indicates a faster mass transfer from the solid phase, corresponding to enhanced mass transfer.  Over time, as the amount of solute decreases, it becomes a limiting factor, reducing the effect of pressure changes on the system. Eventually, the sensitivities approach zero asymptotically as the solute is washed out from the bed.

@@

	Figure \ref{fig:Sensitivty_P_CF} illustrates the sensitivity of solute concentration in the fluid phase to pressure changes. Compared to Figure \ref{fig:Sensitivty_P_CS}, the dynamic behaviour of fluid phase sensitivities is more apparent. Due to advection, the sensitivities related to the fluid phase moves across the system in the direction of flow. Initially, the system response is low despite the pressure increase improving mass transfer, reflecting the previously discussed idle period. As the process continues, the sensitivities increase, indicating faster mass transfer from the solid phase. The corresponding improvement in the mass transfer from the solid particles to the fluid is represented by the positive values in Figure \ref{fig:Sensitivty_P_CF}. When the solute in the solid phase becomes a limiting factor, the extraction rate slows, and sensitivities eventually approaching zero asymptotically.
